Modals across experiments

### 2021-11-11 analysis/exploratory/modality/rscripts/graphs.Rmd
added initial files from old repo

---
title: "Exploratory analysis of context: KL Divergence"
author: Morgan Moyer
date: May 18, 2021
output: html_document
---

@@

# sanity check the join
nrow(context_root) + nrow(context_emb) + nrow(nocontext_root) + nrow(nocontext_emb)
nrow(d)
```

# Calculate KL-Divergence

KLD(P || Q) = sum x in X: (P(x) * log2(P(x)/Q(x)))

Compare to the uniform distribution.

```{r widen the data, warning=FALSE, echo=FALSE, message=FALSE}
# Aggregate ratings across subject and convert to wide
d.wide.context <- d %>%
  filter(Task == "context") %>%
  group_by(tgrep_id,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating)) %>%
  select(tgrep_id,paraphrase,mean_rating) %>%
  pivot_wider(names_from=paraphrase,values_from=mean_rating)

d.wide.nocontext <- d %>%
  filter(Task == "nocontext") %>%
  group_by(tgrep_id,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating)) %>%
  select(tgrep_id,paraphrase,mean_rating) %>%
  pivot_wider(names_from=paraphrase,values_from=mean_rating)
```

```{r calculate kld, warning=FALSE, echo=FALSE, message=FALSE}
kls.context <- d.wide.context %>%
  group_by(tgrep_id) %>%
  summarize(kld = KL(rbind(c(a,every,the),c(1/3,1/3,1/3)),unit="log2"))
kls.context$Task <- "context"

kls.nocontext <- d.wide.nocontext %>%
  group_by(tgrep_id) %>%
  summarize(kld = KL(rbind(c(a,every,the),c(1/3,1/3,1/3)),unit="log2"))
kls.nocontext$Task <- "nocontext"
```

```{r combine together, warning=FALSE, echo=FALSE, message=FALSE}
d.kls <- full_join(kls.nocontext,kls.context)
```

```{r factors, warning=FALSE, echo=FALSE, message=FALSE}
# look at linguistic factors
d.factors <- unique(d[,c("tgrep_id","ModalPresent","Wh","QuestionType")])
kls.factors <- left_join(d.kls,d.factors,by="tgrep_id")
```

## Plot the KLD between context and no context

### Overall
```{r, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = d.kls, aes(x=Task, y = kld, fill=Task)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("graphs/denisty_kld_context.pdf",width=6,height=3)
```

### Factors
```{r, warning=FALSE, echo=FALSE, message=FALSE}
agr <- kls.factors %>% 
  group_by(Task, ModalPresent,Wh) %>% 
  summarize(mean_kld = mean(kld), 
            CILow = ci.low(kld), 
            CIHigh = ci.high(kld)) %>%
  mutate(YMin = mean_kld - CILow, 
         YMax = mean_kld + CIHigh) %>%
  drop_na()

# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))

ggplot(data = agr, aes(x = Task, y=mean_kld, fill=ModalPresent)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),
                width=.25,
                position="dodge", 
                show.legend = FALSE) +
  facet_wrap(~Wh,ncol=2) +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette) +
  scale_alpha_discrete(range = c(.5,1))
# ggsave("graphs/denisty_kld_context_factors.pdf",width=6,height=6)
```

## Difference in KLD by factors
```{r, warning=FALSE, echo=FALSE, message=FALSE}
kls.diffs.factors <- left_join(kls.difs.rel,d.factors,by="tgrep_id")

ggplot(data = kls.diffs.factors, aes(x = difference_in_kld, fill=ModalPresent)) +
  geom_density(alpha = .4) +
  facet_wrap(~Wh) +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("graphs/difference_in_kld_factors_density.pdf",width=10,height=8)

#plot all the individuals mean chosen
ggplot(kls.diffs.factors, aes(x=Wh,y=difference_in_kld,fill=ModalPresent)) +
  geom_bar(stat="identity",position="dodge")
# ggsave("graphs/difference_in_kld_by-item_factors.pdf",width=10,height=8)

@@

## Does Task predict KLD?

- Yes: going from Context to NoContext yields a decrease in KLD
- Interpretation: There is less information loss in the Context Task than in the NoContext Task
```{r, warning=FALSE, echo=FALSE, message=FALSE}
# just recode the factors so that negative values are nocontext and positive values are context
d.kls$Task <- as.factor(d.kls$Task)
d.kls$Task <- recode_factor(d.kls$Task, context=1, nocontext=0)

d.kls$tgrep_id = as.factor(d.kls$tgrep_id)
d.kls$cTask = as.numeric(as.factor(d.kls$Task)) - mean(as.numeric(as.factor(d.kls$Task)))

m.kld <- lmer(kld~cTask + (1|tgrep_id), data=d.kls)
summary(m.kld)

plot(m.kld)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}

d.kls$Fitted = fitted(m.kld)
ggplot(d.kls, aes(x=Fitted,y=kld)) +
  geom_point() +
  geom_smooth(method="lm") +
  xlim(0,1) +
  ylim(0,1) +
  ylab("Empirical rating") +
  xlab("Predicted rating")
# ggsave("../graphs/model_fit_0.pdf",width=5,height=4)
cor(d.kls$Fitted,d.kls$kld)
```

>>>>

---
title: "Modals across experiments"
author: "morgan moyer"
date: "11/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

# sanity check the join
nrow(context_root) + nrow(context_emb) + nrow(nocontext_root) + nrow(nocontext_emb)
names(d)
```

# Let's graph overall
```{r agregate for graphing}
agr = d %>%
  # look only at embedded questions
  # filter(VerbLemma != "") %>% 
  group_by(Modal,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
dodge = position_dodge(.9)
```

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  facet_wrap(~Modal) +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  # guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette)
    # scale_alpha_discrete(range = c(.5,1))
# ggsave("../graphs/overall_negation.pdf",width=6,height=3)

@@

# Let's graph with Task
```{r agregate for graphing}
agr = d %>%
  # look only at embedded questions
  # filter(VerbLemma != "") %>% 
  group_by(Task,Modal,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
dodge = position_dodge(.9)
```

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  facet_grid(Task~Modal) +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  # guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette)
    # scale_alpha_discrete(range = c(.5,1))
# ggsave("../graphs/overall_negation.pdf",width=6,height=3)
